---
output: 
  html_document:
    theme: yeti
    css: style.css
    highlight: tango
---

```{r setup, include=FALSE, eval=FALSE}
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
```


<center><h1><b>RICH A. PAULOO</b></h1></center>
***  
<!-- <center> -->
<!--   <i class="fa fa-globe" aria-hidden="true"></i> [richpauloo.github.io](https://richpauloo.github.io/) | -->
<!--   <i class="fa fa-github" aria-hidden="true"></i> [richpauloo](https://github.com/richpauloo) | -->
<!--   <i class="fa fa-envelope" aria-hidden="true"></i> rpauloo@ucdavis.edu -->
<!-- </center> -->

<!-- font awesome hack to make doc < 2MB in size -->
<center>
  <img src = "web.png" height = 15, width = 15> [richpauloo.github.io](https://richpauloo.github.io/) |
  <img src = "git.png" height = 15, width = 15> [richpauloo](https://github.com/richpauloo) |
  <img src = "mail.png" height = 15, width = 15> rpauloo@ucdavis.edu 
  <!-- <img src = "home.png" height = 15, width = 15> 740 Miller Drive. Davis, CA 95616 -->
</center>

***  

```{r, echo = FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(r2d3)
r2d3(data = c(), d3_version = 4, script = "voronoi.js", width = '100%', height = 100)
```

<!-- **OBJECTIVE**: Data Scientist Intern, PhD, 2019 - Mountain View, New York or Seattle -->

**Hydrogeology PhD (March 2020)** and domain expert in hydrologic sciences skilled in the development of physically-based, 3D regional groundwater flow and contaminant transport models. Demonstrated ability to work in multidisciplinary teams on projects bridging physical simulation, climate science, and disadvantaged populations, with success in two data science competitions hosted by NASA and the West Big Data Hub. Possess 3+ years scripting non-routine, end-to-end quantitative analysis of spatial, tabular, natural language, and timeseries data in SQL, R, and Python. Creative thinker skilled in research design, from data gathering and processing, to statistics, machine learning, data visualization, communication, and drawing policy implications from research. Possess strong interpersonal, presentation, writing, and oratory skills honed through presentations to scientists, policymakers, and the general public. Able to grasp new concepts and ideas quickly. Motivated by learning new skills and using science to make the world a safer, cleaner, more equitable place. 

***

## EDUCATION

**University of California, Davis, *Ph.D. in Hydrogeophysics.*** \hfill (March 2020)  
<ul> 
Dissertation: *"Linking Water Quality and Quantity Objectives: towards integrated, sustainable, basin-scale groundwater management"* (Advised by Dr. Graham E. Fogg)
</ul>
    
**University of California, Berkeley, *B.S. in Integrative Biology*** \hfill (2011)  
<ul> 
*minor in Conflict Resolution*
</ul>

***  


## EMPLOYMENT

<p class="alignleft">**Graduate Student Researcher at University of California, Davis**</p><p class="alignright">2015--present</p>  

<br>

<ul> 
  <li>As a **Researcher** for the [U.S./China Clean Energy Research Center](https://cerc-wet.berkeley.edu/), to address the long-term sustainability of fresh groundwater resources in the US and China, I developed novel, computationally efficient, regional-scale physical simulation models to assess basin-scale groundwater quality degradation due to non-point source pollution. These models are the first of their kind, allowing efficient simulation of basin-scale 3D contaminant transport processes using 2D equivalent models. I used R to prepare and automate simulations in MODFLOW, and ParaView for visualization and LaTeX for reporting.</li>  
  <li>As a **National Science Foundation Fellow** in the Climate Change, Water, and Society interdisciplinary graduate education and research traineeship, I aimed to quantify the impact of drought on vulnerable populations in California. I led a team consisting of a hydrologist (myself), economist, engineer, statistician, and social scientist to develop spatial, statistical, and machine learning models to predict domestic well failure in California during the 2012-2016 drought and in future drought scenarios. We found that drought disproportionally hurt vulnerable, low-income communities, and developed policy recommendations for meeting these challenges. We won the statewide 2018 [California Water data Challenge](http://waterchallenge.data.ca.gov/) for our work, [which can be viewed here](https://richpauloo.github.io/flexdash.html). I also collaborated with climate scientists, social scientists, and economists to co-organize and convene [a conference](https://watershed.ucdavis.edu/news/2016/03/09/april-4-5-weathering-change-impact-climate-change-and-sgma-ca-water-conference) on critical issues at food-water-energy-nexus.</li>
  <li>As a **Data Engineer and Developer** for UC Water, I tackled the need for free and open-source groundwater monitoring tools, and the lack of access to high-quality lithologic data. I completed two open-source, web-based tools to meet these needs: (1) a [groundwater monitoring web-dashboard](http://ucwater.org/gw_obs/) built with R, SQL, Shiny, and Shiny Server that includes: automated end-to-end extraction and analysis of telemetry-based data, automated report generation, and interactive data visualization. (2) a [spatial database query tool](http://ucwater.org/oswcr/) that allows users to query a cleaned version of the California well completion report database using a shapefile defining their area of interest. My data has since been used and cited by consultants, academic researchers, and non-profit think tanks.</li>  
  <li>As a **Data Science Consultant** for the [UC Davis Data Science Initiative](http://dsi.ucdavis.edu/), to understand CIG software package interactions in academic journals, I used R for text extraction, natural language processing, and network analysis of ~500 journal articles, showing how usage has changed over time and the relationships between packages.</li>  
  <li>I developed and maintain the [Hydrologic Sciences Graduate Group webpage](https://hsgg.ucdavis.edu/).</li>  
</ul>
 
 
<p class="alignleft">**Expedition Leader at National Geographic**</p><p class="alignright">2016</p> 

<br>

<ul> 
  <li>Based out of various locations in Thailand, and in close collaboration with Thai program partners, I co-coordinated global cultural awareness activities, interpersonal skill development, emotional intelligence, teambuilding, reflection, and community service activities for a group of 16 international high-school students on a 14-day expedition.</li>  
</ul>


<p class="alignleft">**Site Manager at NatureBridge**</p><p class="alignright">2014-2016</p> 

<br>

<ul> 
  <li>Access to environmental education is critical to inspiring future generations. At the educational 501(c)(3) NatureBridge, I managed and trained new staff, practiced conflict resolution, and coordinated with the Operations team to run a campus of 200 hundreds students and 60 employees in the Marin Headlands. I analyzed the budget to evaluate the feasibility of an employee meal program, designed a study to measure collaboration among educators, distributed my findings to management and staff in reports, and piloted an employee meal program that increased instructor collaboration interactions by 200%.</li>
</ul>


<p class="alignleft">**Environmental Science Teacher at San Mateo Outdoor Education**</p><p class="alignright">2012-2013</p> 

<br>

<ul> 
  <li>Worked with diverse groups of students, parents, and teachers to facilitate interpersonal skill development, emotional intelligence, teambuilding, and environmental science curriculum to large groups (upwards of 200) of 5th and 6th graders in an outdoor education setting. Extensive lesson planning and curriculum development, nonviolent communication, co-teaching, and conflict resolution between students.</li>  
</ul>





*** 


## GRANTS & AWARDS

<ul>

  <li><p class="alignleft">**($111,000)** *NSF-IGERT in Climate Change, Water, and Society*, National Science Foundation</p><p class="alignright">2015</p></li><br>
  
  <li><p class="alignleft">*NSF-GRFP Honorable Mention*, National Science Foundation</p><p class="alignright">2016</p></li><br>
  
  <li><p class="alignleft"> **($1,500)** *2018 California Water Data Challenge Winner* (Statewide Competition). [[**VIEW PROJECT**](https://richpauloo.github.io/flexdash.html)]</p><p class="alignright">2018</p></li><br>
  
  <li><p class="alignleft">**($1,400)** *NASA Data Visualization Competition Runner Up* (National Competition). [[**VIEW PROJECT**](https://github.com/richpauloo/richpauloo.github.io/blob/master/presentations/pauloo_hyperwall_no_lines.pdf)]</p><p class="alignright">2018</p></li><br>
  
  <li><p class="alignleft">**($1,000)** *Early Career Data Science Researcher Award*, West Big Data Hub</p><p class="alignright">2018</p></li><br>
  
</ul>



***

## LANGUAGES

* Spanish (conversational)  
* Portuguese (elementary)  

*** 


## VOLUNTEERING, COMMUNICATION, & OUTREACH

* Reviewer for [Water Resources Research](https://agupubs.onlinelibrary.wiley.com/journal/19447973)  
* Personal Research and Data Science Blog (**~ 1.5k monthly users**) [[**VIEW BLOG**]](https://richpauloo.github.io/). Example topics include:  
    + [Chi-squared tests in R](https://richpauloo.github.io/2018-02-03-Tidy-Chi-Squared-stats-in-infer/)  
    + [predicting fatal landslides worldwide with machine learning](https://richpauloo.github.io/landslides.html)  
    + [data science software installation guides](https://richpauloo.github.io/2018-05-16-Installing-the-R-kernel-in-Jupyter-Lab/)  
* [*Eastern San Joaquin Valley and other CA drinking water supplies at risk in the next drought*](https://californiawaterblog.com/2018/11/12/domestic-well-vulnerability-to-drought-in-californias-central-valley/), California Water Blog  

***

## PRESENTATIONS

**2019. Pauloo, R., G.E. Fogg, R., Escriva-Bou, A., Fencl, A., Guillon, H.** "An Overview of Domestic Well Data in California's Central Valley: Opportunities for Informed Risk Assessment." California Environmental Protection Agency. Sacramento, CA. 18 Jan. [[**VIEW SLIDES**]](https://github.com/richpauloo/richpauloo.github.io/blob/master/presentations/2018_cal_epa.pdf)

**2018. Pauloo, R., Fogg, G.E., Z. Guo, C.V. Henri** "Upscaling Regional, 3D, Non-Fickian Solute Transport with 2D Equivalent Models." American Geophysical Union. Final Paper Number H43I-2582. Washington, DC. 13 Dec. [[**VIEW POSTER**]](https://github.com/richpauloo/richpauloo.github.io/blob/master/presentations/agu_pauloo_2018.pdf)  

**2018, Pauloo, R.** "An Interactive Mapping Interface to the California Online State Well Completion Report Database" UC Water Annual Meeting. Sacramento, CA. 25 Oct. [[**VIEW PRESENTATION**]](https://shiny.lawr.ucdavis.edu/shiny/oswcrd/)  

**2018, Pauloo, R., A. Escriva-Bou, A. Fencl, H. Guillon** "Domestic Well Vulnerability to Drought in California's Central Valley" California Water Data Challenge Awards Summit. Los Angeles, CA. 18 Oct.  

**2018. Pauloo, R., A. Escriva-Bou, A. Fencl, H. Guillon** "Data-Driven Prediction of Domestic Well Failure in California, and Forecasting Future Well Failure with Machine Learning Models" CA Water Data Summit. Los Angeles, CA. 23 Sept.  

**2018. Pauloo, R.** "Regional Groundwater Quantity and Quality Models" Stanford University. Palo Alto, CA. 24 May.  

**2017. Pauloo, R., Fogg, G.E., Z. Guo, C.V. Henri** "Basin Scale Groundwater Quality Management Modeling: Methods to Upscale the Advection Dispersion Equation." American Geophysical Union. Final Paper Number H51G-1360. New Orleans, LA. 14 Dec. [[**VIEW POSTER**]](https://github.com/richpauloo/richpauloo.github.io/blob/master/presentations/agu_pauloo_2017_small.pdf)  

**2017. Pauloo, R., Fogg, G.e., E. Goharian, J.D. Herman** "How to Maximize Total Water Storage in California." UC Water Annual Retreat. Davis, CA. 21--22 Sept. 

**2016. Pauloo, R., Fogg, G.E., Z. Guo** "Water Resource Vulnerability to Groundwater Salinization in Closed Basin Aquifers." American Geophysical Union. Final Paper Number H33F-1610. San Francisco, CA. 14 Dec. 

**2016. Pauloo, R., Fogg, G.E., Z. Guo** "Investigating Groundwater Sustainability and Water Quality Degradation in California's Tulare Basin using a Simple Mixing Model." Groundwater Resources Association Annual Meeting. Concord, CA. 28 Sept.

**2016. PPauloo, R., Fogg, G.E., Z. Guo** "Groundwater Salinization due to Hydraulic Closure in Tulare Basin over long time scales." Toward Sustainable Groundwater in Agriculture. Burlingame, CA. 28 June.

**2015. Pauloo, R., P. Marielle** "The Past Present and Future: Long-term climate trends and water management history in California." Water Scarcity in the West: Past Present and Future. Davis, CA. 12--13 April. [[**VIEW POSTER**]](https://github.com/richpauloo/richpauloo.github.io/blob/master/presentations/2017_04_18_weathering_change_poster.pdf)  


*** 

## PUBLICATIONS

<!-- ### **Journal Articles** -->

**Pauloo, R.**, A. Escriva-Bou, A. Fencl, H. Guillon 2019. "Domestic Well Vulnerability to Drought Duration and Unsustainable Groundwater Management in California's Central Valley." *PNAS* *(in preparation)*

**Pauloo, R.**, C. Henri, Z. Guo, G.E. Fogg 2019. "Upscaling Regional, 3D, Non-Fickian Solute Transport with 2D Equivalent Models." *Water Resources Research* *(in preparation)*

**Pauloo, R.**, Z. Guo, G.E. Fogg 2019. "Groundwater Development Induced Basin Closure: An Unrecognized Threat to Water Quality Sustainability." *Advances in Water Resources* *(in preparation)*

*** 



 
<!-- font awesome hack to make doc < 2MB in size -->

<!-- <center> -->
<!-- <h5>Made with <img src = "heart.png" height = 15, width = 15> &  <img src = "coffee.png" height = 15, width = 15>  by Rich A. Pauloo</h5> -->
<!-- </center> -->

<!-- <center> -->
<!-- <h5>Made with <i class="fa fa-heart" aria-hidden="true"></i> & <i class="fa fa-coffee" aria-hidden="true"></i> by Rich A. Pauloo</h5> -->
<!-- </center> -->

<center>
  <img src = "web.png" height = 15, width = 15> [richpauloo.github.io](https://richpauloo.github.io/) |
  <img src = "git.png" height = 15, width = 15> [richpauloo](https://github.com/richpauloo) |
  <img src = "mail.png" height = 15, width = 15> rpauloo@ucdavis.edu 
</center>

<!-- <center> -->
<!--   <i class="fa fa-globe" aria-hidden="true"></i> [richpauloo.github.io](https://richpauloo.github.io/) | -->
<!--   <i class="fa fa-github" aria-hidden="true"></i> [richpauloo](https://github.com/richpauloo) | -->
<!--   <i class="fa fa-envelope" aria-hidden="true"></i> rpauloo@ucdavis.edu -->
<!-- </center> -->

<br>
<br>


